Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-09-11 temp/VHzQ_pre2019/WW4C_VHzQ_20180817.tex
tex/older_tex/ removed

\title[Very high-$z$ Quasars]
        {Near and mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
       {Nicholas P. Ross$^{1}$\thanks{Corresponding Author: npross@roe.ac.uk} and Nicholas J. G. Cross$^{1}$
\\ 
$^1$Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\
}

\maketitle
\begin{abstract}
Very high redshift quasars are excellent probes of physics the early
Universe.  Understanding how billion solar mass black holes
formed, evolve and accrete in the first billion years is a key
outstanding problem in contemporary astrophysics, and one that wil lbe
directly addressed by upcoming missions such as ESA {\it Euclid} and
the {\it James Webb Space Telescope}.
%%
We can gain a good appreciation for what these missions will discover
by collating the datasets we currently have, and in this paper, we,
for the first time since the discovery of $z\geq5.00$ quasars assemble all spectroscopically confirmed very high redshift quasars in
one catalogue.
%%
In particular we present the near ($z/Z, y/Y, J, H, K_{s}$ and $K$)
infrared and mid-infrared (WISE) properties of all 424
spectroscopically known redshift $z\geq5$ quasars.  Using archival public
WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric variability
in the near-infrared and find blah.  We present a comprehensive series
of colour-redshift and colour-colour plots and make inferences into
the hot dust properties of the very high-redshift quasar
population. Extrapolating the known quasar luminosity function we
suggest that 80\% of the possibly detected $z\geq5$ quasars in the
current datasets have been discovered.
\end{abstract}

@@

\iffalse
\section*{TO DOs}
{\bf For NPR: }
\begin{itemize}
\item Almost certainly want to compile M$_{\rm BH}$; hmmmm....
\item May well wanna try and get M$_{\star}$ too...!! (yuck); 
\item Filter curve plot with LSST, Euclid and even wide JWST filters...; 
\item Check with SpIES/SHELA (but only 3.4/4.6$\mu$m; 
\item Concentrate on NIR and W1/W2
\item SED dust plot with just W1/W2...
\item Check vs. Assef R90/C75 catalog.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Very high redshift quasars (VH$z$Q; and defined here to have redshifts
$z\geq5.00$) are excellent probes of the early Universe. This includes
studies of the Epoch of Reionization for hydrogen \citep[see e.g.][for
reviews]{Fan2006review, Mortlock2016}, the formation and build-up of
supermassive black holes \citep[e.g., ][]{Rees1984, WyitheLoeb2003,
Volonteri2010, Agarwal2016, Valiante2018, Latif2018} and early metal
enrichment \citep[see e.g., ][]{Simcoe2012, Chen2017, Bosman2017}.

Super-Eddington (also called super-critical) accretion is a viable
physical mechanism to explain the high luminosity and rapid growth of
supermassive black holes in the early universe
\citep[e.g.,][]{AlexanderNatarajan2014, MadauHaardtDotti2014,
Volonteri2015, Pezzulli2016, Lupi2016} Thus, one might well expect
VH$z$Qs to vary in luminosity as they potentially go through phases of
Super-Eddington accretion and these should be looked for, especially
when the e.g.  near-infrared $K$-band is sampling rest-frame
$\approx$3670, 3145, 2750\AA\ at redshifts $z=5, 6$ and 7,
respectively.

The build-up of SMBHs via accretion is important due to the desire to
understand the evolution of the black hole mass function (BHMF) and
since there is a remarkable similarity between cosmic star-formation
rate density (SFRD) and black hole accretion rate (BHAR). The SFRD and
BHAR trace each other \citep[with a normalisation factor of
$\sim3000$,][]{Willott2013b, MadauDickinson2014} across redshifts
$z=0-4$.  However, new results \citep[e.g., ][]{Vito2018a, Calhau2018}
suggest that the SFRD and BHAR do {\it not} have the same evolutionary form
at higher redshift, $z>4$. The physical reasons for this are currently very poorly
understood.  For the high-$z$ QSOs that we can measure BHAR and SFR, most
have BHAR $\gg 10^{-3}$ SFR, indeed {\it Herschel}-detected sources at
$z\simeq.8$ have $L_{\rm SF} \simeq L_{\rm AGN}$ \citep[e.g.,
][]{Netzer2014} and {\it Herschel}-detected quasars show no evolution
in the $M_{\rm BH} - M_{\star}$ relation from $z\sim2$ to $z\sim0$.
where $M_{\rm BH}, M_{\rm bulge}, M_{\star}$ are the SMBH mass,
stellar mass of the bulge and galaxy total stellar mass,
respectively. The evolution and potential importance of the SFRD and BHAR 
link is a key outstanding problem in contemporary astrophysics, 
potentially explaining the galaxy-black hole scaling relations. 
However, there may also simply be a non-causal origin of the black-hole 
galaxy scaling relations \citep{Jahnke2011}.

Quasars are also known to be prodigious emitters of infrared emission,
thought to be from the thermal emission of dust grains heated by
continuum emission from the accretion disc \citep[][]{Hill2014,
Hickox2017}.  Observations in the mid-infrared, e.g. $\sim$3-30$\mu$m
allow discrimination between AGN\footnote{Historically, ``quasars'' and
``Active Galactic Nuclei (AGN)'' have described different
luminosity/classes of objects, but here we use these terms
interchangeably (with a preference for quasar) in recognition of the
fact that they both describe accreting supermassive black holes
\citep[e.g.][]{Haardt2016book}.}  and passive galaxies due to the
1.6$\mu$m ``bump'' entering the MIR at $z\approx0.8-0.9$ \citep[e.g.,
][]{Wright1994, Sawicki2002, Lacy2004, Stern2005, Richards2006b,
Timlin2016} as well as AGN and star-forming galaxies due to the
presence of Polycyclic Aromatic Hydrocarbon (PAHs) at $\lambda >3\mu$m
\citep[e.g., ][]{Yan2007, Tielens2008}.

@@

We chose redshift $z=5.00$ as our lower redshift limit due to a combination 
of garnishing a large sample, adequately spanning physical properties 
(e.g. luminosity, age of the Universe) and to incorporate what knowledge 
we have gained over the last couple of decades since $z>5$ quasars were 
discovered.

This paper can be considered an update of \citet{Blain2013} and also
an extension of parts of \citet{Banados2016}, with the latter study reporting
WISE W1, W2, W3 and W4 magnitudes for Panoramic Survey Telescope and Rapid Response System 1 
\citep[Pan-STARRS1, PS1;][]{Kaiser2002, Kaiser2010}, but with no  
further investigation into the reddest WISE waveband for the VH$z$Qs.
%%
\citet{Banados2016} reports and investigates the W1, W2 and W3 properties
of quasars at $z > 5.6$.

@@

Specifically, we use data from: \citet{Fan2000}, \citet{Fan2001c},
\citet{Fan2003} , \citet{Fan2004}, \citet{Mahabal2005},
\citet{Cool2006}, \citet{Fan2006}, \citet{Goto2006},
\citet{McGreer2006}, \citet{Carilli2007}, \citet{Kurk2007},
\citet{Stern2007}, \citet{Venemans2007}, \citet{Willott2007},
\citet{Jiang2008}, \citet{Wang2008}, \citet{Jiang2009}, \citet{Kurk2009}, 
\citet{Mortlock2009}, \citet{Willott2009}, \citet{Carilli2010}, 
\citet{Wang2010}, \citet{Willott2010a}, \citet{Willott2010b}, 
\citet{DeRosa2011}, \citet{Mortlock2011}, \citet{Wang2011}, 
\citet{Zeimann2011}, \citet{Morganson2012}, \citet{Venemans2012}, 
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Wang2013}, 
\citet{Willott2013b}, \citet{Banados2014}, \citet{Calura2014}, 
\citet{Leipski2014}, \citet{Banados2015a}, \citet{Banados2015b}, 
\citet{Becker2015}, \citet{Carnall2015}, \citet{Jiang2015}, 
\citet{Kashikawa2015}, \citet{Kim2015}, \citet{Reed2015}, 
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Willott2015}, 
\citet{Wu2015}, \citet{Venemans2016}, \citet{Wang2016_WISE}, 
\citet{Matsuoka2016}, \citet{WangR2016}, \citet{Mortlock2011},
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Venemans2013},
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Banados2016},
\citet{Matsuoka2016}, \citet{Reed2017}, \citet{Wang2017},
\citet{Mazzucchelli2017}, \citet{Ikeda2017}, \citet{Tang2017},
\citet{Koptelova2017}, \citet{Banados2018}, \citet{Matsuoka2018a} 
and \citet{Matsuoka2018b}.

@@

@@

\section{A. Photometric Bands and Conversions}
    Due to the differing normalizations between the
    SDSS and  UKIDSS photometric systems, certain corrections are required.  To present
    our data in the  purest sense, all the NIR magnitudes from UKIDSS
    (originally AB magnitudes)  were corrected to Vega magnitudes as
    suggested in \citet{Hewett06}.

@@

\begin{table*}[h*]

@@

   \caption{Adapted from Table 9 of \citet{Peth11}. 

@@

@@

@@

@@

      $Y_{\rm VISTA}$     &    10184    &   9427      &   10977   &    905        & $Y_{\rm VISTA} $     &$ = Y_{AB}  - 0.601 $          \\

@@

      $J_{\rm VISTA} $     &     10184   &      9427   &    10977   &     905     &  $J_{\rm VISTA}  $     & $= J_{AB}    - 0.938  $         \\

@@

      $H_{\rm VISTA}$      &    16310    &    14604   &    18422   &   2833     & $H_{\rm VISTA}$      & $= H_{AB}  - 1.368 $        \\
      $K$s$_{\rm VISTA}$     &    21337    &    19333  &    23674   &   3055     & $K$s$_{\rm VISTA}$      & $ = K_{AB} - 1.83  $          \\ 

>>>>

\title[Very high-$z$ Quasars]
        {Near and Mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
       {Nicholas P. Ross$^{1}$\thanks{Corresponding Author: npross@roe.ac.uk} and Nicholas J. G. Cross$^{1}$
\\ 
$^1$Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\
}

\maketitle
\begin{abstract}
In this paper, we, for the first time since the discovery of $z\geq5$
quasars, assemble all spectroscopically confirmed very high redshift
quasars in one catalogue.  In particular we present the near
($zZyYJHK_{s}$ and $K$) infrared and mid-infrared (WISE) properties of
all 424 spectroscopically confirmed redshift $z\geq5.00$ quasars.
Using archival public WFCAM/UKIRT and VIRCAM/VISTA data we check for
photometric variability in the near-infrared that might be expected
from Super-Eddington accretion and find {\it blah}.  We present a
comprehensive series of colour-redshift and colour-colour plots and
make inferences into the hot dust properties of the very high-redshift
quasar population. Extrapolating the known quasar luminosity function
we suggest that $x$\% of the possibly detected $z\geq5$ quasars in the
current datasets have been discovered.
\end{abstract}

@@

\iffalse
\section*{TO DOs}
{\bf For NPR: }
\begin{itemize}
\item Concentrate on reporting NIR and W1/W2 figures and tables
\item SED dust plots 
\item Check vs. Assef R90/C75 catalog.
\item Filter curve plot with LSST, Euclid and even wide JWST filters...; 
\item Check with SpIES/SHELA (but only 3.4/4.6$\mu$m; 
\item Almost certainly want to compile M$_{\rm BH}$; THIS CAN WAIT, otherwise is massive paper-creep
\item May well wanna try and get M$_{\star}$ too...!! (yuck); THIS CAN WAIT, otherwise is massive paper-creep
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Very high redshift quasars (VH$z$Q; defined here to have redshifts
$z\geq5.00$) are excellent probes of the early Universe. This includes
studies of the Epoch of Reionization for hydrogen \citep[see e.g.][for
reviews]{Fan2006review, Mortlock2016}, the formation and build-up of
supermassive black holes \citep[e.g., ][]{Rees1984, WyitheLoeb2003,
Volonteri2010, Agarwal2016, Valiante2018, Latif2018} and early metal
enrichment \citep[see e.g., ][]{Simcoe2012, Chen2017, Bosman2017}.

Super-Eddington (also called super-critical) accretion is a viable
physical mechanism to explain the high luminosity and rapid growth of
supermassive black holes in the early universe
\citep[e.g.,][]{AlexanderNatarajan2014, MadauHaardtDotti2014,
Volonteri2015, Pezzulli2016, Lupi2016, Pezzulli2017, Takeo2018}. Thus,
one might well expect VH$z$Qs to vary in luminosity as they
potentially go through phases of Super-Eddington accretion and these
signatures of photometric variablity should be looked for, especially
when the e.g. near-infrared $K$-band is sampling rest-frame
$\approx$3670, 3145, 2750\AA\ at redshifts $z=5, 6$ and 7,
respectively.

Quasars are also known to be prodigious emitters of infrared emission,
thought to be from the thermal emission of dust grains heated by
continuum emission from the accretion disc
\citep[e.g.,][]{Richards2006b, Leipski2014, Hill2014, Hickox2017}.
Observations in the mid-infrared, e.g. $\sim$3-30$\mu$m allow
discrimination between AGN\footnote{Historically, ``quasars'' and
``Active Galactic Nuclei (AGN)'' have described different
luminosity/classes of objects, but here we use these terms
interchangeably (with a preference for quasar) in recognition of the
fact that they both describe accreting supermassive black holes
\citep[e.g.][]{Haardt2016book}.}  and passive galaxies due to the
1.6$\mu$m ``bump'' entering the MIR at $z\approx0.8-0.9$ \citep[e.g.,
][]{Wright1994, Sawicki2002, Lacy2004, Stern2005, Richards2006b,
Timlin2016} as well as AGN and star-forming galaxies due to the
presence of Polycyclic Aromatic Hydrocarbon (PAHs) at $\lambda >3\mu$m
\citep[e.g., ][]{Yan2007, Tielens2008}.

@@

This paper can be considered an update of \citet{Blain2013} and also
an extension of parts of \citet{Banados2016}, with the latter study
reporting WISE W1, W2, W3 and W4 magnitudes for Panoramic Survey
Telescope and Rapid Response System 1 \citep[Pan-STARRS1,
PS1;][]{Kaiser2002, Kaiser2010}, but with no further investigation
into the reddest WISE waveband for the VH$z$Qs.  \citet{Banados2016}
reports and investigates the W1, W2 and W3 properties of quasars at $z
> 5.6$. We chose redshift $z=5.00$ as our lower redshift limit due to
a combination of garnishing a large sample, adequately spanning
physical properties (e.g. luminosity, age of the Universe) and to
incorporate what knowledge we have gained over the last couple of
decades since $z>5$ quasars were discovered.

@@

Specifically, we use data from: \citet{Fan2000}, \citet{Fan2001c},
\citet{Fan2003}, \citet{Fan2004}, \citet{Mahabal2005},
\citet{Cool2006}, \citet{Fan2006}, \citet{Goto2006},
\citet{McGreer2006}, \citet{Carilli2007}, \citet{Kurk2007},
\citet{Stern2007}, \citet{Venemans2007}, \citet{Willott2007},
\citet{Jiang2008}, \citet{Wang2008}, \citet{Jiang2009}, \citet{Kurk2009}, 
\citet{Mortlock2009}, \citet{Willott2009}, \citet{Carilli2010}, 
\citet{Wang2010}, \citet{Willott2010a}, \citet{Willott2010b}, 
\citet{DeRosa2011}, \citet{Mortlock2011}, \citet{Wang2011}, 
\citet{Zeimann2011}, \citet{Morganson2012}, \citet{Venemans2012}, 
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Wang2013}, 
\citet{Willott2013b}, \citet{Banados2014}, \citet{Calura2014}, 
\citet{Leipski2014}, \citet{Banados2015a}, \citet{Banados2015b}, 
\citet{Becker2015}, \citet{Carnall2015}, \citet{Jiang2015}, 
\citet{Kashikawa2015}, \citet{Kim2015}, \citet{Reed2015}, 
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Willott2015}, 
\citet{Wu2015}, \citet{Venemans2016}, \citet{Wang2016_WISE}, 
\citet{Matsuoka2016}, \citet{WangR2016}, \citet{Mortlock2011},
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Venemans2013},
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Banados2016},
\citet{Matsuoka2016}, \citet{Reed2017}, \citet{Wang2017},
\citet{Mazzucchelli2017}, \citet{Ikeda2017}, \citet{Tang2017},
\citet{Koptelova2017}, \citet{Banados2018}, \citet{Matsuoka2018a} 
and \citet{Matsuoka2018b}.

@@

%%
We can gain a good appreciation for what these missions will discover
by collating the datasets we currently have.

@@

\section{A. Photometric Bands and Conversions}
    Due to the differing normalizations between the
    SDSS and  UKIDSS photometric systems, certain corrections are required.  To present
    our data in the  purest sense, all the NIR magnitudes from UKIDSS
    (originally AB magnitudes)  were corrected to Vega magnitudes as
    suggested in \citet{Hewett2006}.

@@

\begin{table*}

@@

   \caption{Adapted from Table 9 of \citet{Peth2011}. 

@@

%%
%%
From \citet{GonzalezFernandez2018} 
$Z_{\rm AB}   -  Z_{\rm Vega}  = 0.502$;  
$Y_{\rm AB}  -  Y_{\rm Vega}    = 0.600 $;
$J_{\rm AB}   -  J_{\rm Vega}    = 0.916  $;
$H_{\rm AB}  -  H_{\rm Vega}    = 1.366 $;
$Ks_{\rm AB}  -  Ks_{\rm Vega}  = 1.827 $;
%%
and the CASU Vega to AB conversions v1.3:: 
	Z,Y,J,H,Ks were: 0.524, 0.618, 0.937, 1.384, 1.839. 
%%
So, $\Delta$(vs. Gonzalez-Fernandez)::
	(11.2,     1.1,    5.4,     1.6,    0.1) millimags. 
$\Delta$(vsCASU v1.3)::
	(-10.8, -16.9, -15.6,  -16.4, -11.9) millimags. 

@@

      &&&&&&\\

@@

      $Z_{\rm VIRCAM}$  &     8762   & 	 8157    &      9400   &      978    &   $Z_{\rm VIRCAM}  $   &$    = z_{\rm AB} - 0.513 $     \\
      $Z_{\rm WFCAM}$  &     8802   & 	 8129    &      9457   &      926    &   $Z_{\rm WFCAM}  $   &$    = z_{\rm AB} - 0.514 $     \\

@@

      $Y_{\rm VIRCAM}$     &    10184    &   9427      &   10977   &    905        & $Y_{\rm VISTA} $     &$ = Y_{AB}  - 0.601 $          \\
      &&&&&&\\
      $J_{\rm VIRCAM} $     &   12464   &      11427   &    13759   &  1628     &  $J_{\rm VISTA}  $     & $= J_{AB}    - 0.921  $         \\

@@

      &&&&&&\\

@@

      $H_{\rm VIRCAM}$      &    16310    &    14604   &    18422   &   2833     & $H_{\rm VISTA}$      & $= H_{AB}  - 1.368 $        \\
      &&&&&&\\
      $K$s$_{\rm VIRCAM}$     &    21337    &    19333  &    23674   &   3055     & $K$s$_{\rm VISTA}$      & $ = K$s$_{AB} - 1.83  $          \\ 
